---
title: "Datasaurus"
author: "Stephan Huber"
date: "2023-05-16"
output:
  unilur::tutorial_html_solution:
    collapse: TRUE
    solution_suffix: _exercise
  unilur::tutorial_html: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
options(knitr.table.format = "html")
```


a) Load the packages `datasauRus` and `tidyverse`. If necessary, install these packages.

```{r solution=FALSE}
#install.packages("datasauRus")
library(datasauRus)
```

b) The package`datasauRus` comes with a dataset in two different formats: `datasaurus_dozen` and `datasaurus_dozen_wide`. Store them as `ds` and `ds_wide`.  

```{r, solution = FALSE}
ds <- datasaurus_dozen
ds_wide <- datasaurus_dozen_wide
```

c) Open and read the R vignette of the `datasauRus` package. Also open the R documentation of the dataset `datasaurus_dozen`.

```{r, solution = FALSE}
??datasaurus
```

d) Explore the dataset: What are the dimensions of this dataset? Look at the descriptive statistics.

```{r, solution = FALSE}
ds
dim(ds)
head(ds)
glimpse(ds)
view(ds)
summary(ds)
```

e) How many unique values does the variable `dataset` of the tibble `ds` have? Hint: The function unique() return the unique values of a variable and the function length() returns the length of a vector, such as the unique elements.

```{r, solution = TRUE}
unique(ds$dataset)  
unique(ds$dataset) |> 
  length()
```

f) Compute the mean values of the `x` and `y` variables for each entry in `dataset`. Hint: Use the group_by() function to group the data by the appropriate column and then the summarise() function to calculate the mean.


```{r, solution = TRUE}
ds |> 
  group_by(dataset) |> 
  summarise(mean_x = mean(x),
            mean_y = mean(y))
```

g) Compute the standard deviation, the correlation, and the median in the same way. Round the numbers.


```{r, solution = TRUE}
ds |> 
  group_by(dataset) |> 
  summarise(mean_x = round(mean(x),2),
            mean_y = round(mean(y),2),
            sd_x = round(sd(x),2),
            sd_y = round(sd(y),2),
            med_x = round(median(x),2),
            med_y = round(median(y),2),
            cor = round(cor(x,y),4))
```

h) What can you conclude?

```{asis,  solution = TRUE}
The standard deviation, the mean, and the correlation are basically the same for all datasets. The median is different.
```

i) Plot all datasets of `ds`. Hide the legend. Hint: Use the `facet_wrap()` and the `theme()` function.

```{r, fig.height = 8, fig.asp = 1, solution = TRUE}
ggplot(ds, aes(x = x, y = y)) +
  geom_point() +
  facet_wrap(~ dataset, ncol = 3) +
  theme(legend.position = "none") 
```


j) Create a loop that generates separate scatter plots for each unique datatset of the tibble `ds`. Export each graph as a png file.

```{r eval=F, echo=T,  solution=TRUE}
# Get unique values
uni_ds <- unique(ds$dataset)    

# Scatter plot for each dataset
for (uni_v in uni_ds) {
# Select data for the current value
  subset_ds <- ds |> 
    filter(dataset == uni_v) %>%
    select(x, y)
  
  # Make plot
  graph <- ggplot(subset_ds, aes(x = x, y = y)) +
    geom_point() +
    labs(title = paste("Dataset:", uni_v),
         x = "X",
         y = "Y") +
    theme_bw()
  
  # Save the plot as a PNG file
  filename <- paste0("../pic/", "plot_ds_", uni_v, ".png")
  ggsave(filename, plot = graph)
}






