# Learn interactively with _swirl_ {#sec:swirl}
  
The R package _swirl_ makes it fun and easy to learn R programming and data science. 
_swirl_ teaches you R programming and data science interactively, at your own pace, 
and right in the R console! You get immediate feedback on your progress.
If you are new to R, have no fear. _swirl_ will walk you through each of the steps 
required to employ Rstudio and R for your purpose.
To start it, please follow my instructions precisely! 

Open Rstudio and type in the console the following:

```{r eval=FALSE}
install.packages("swirl")
library("swirl")
ls()
rm(list=ls())
install_course_github("hubchev", "swirl-it")
swirl()
``` 


The above lines of code do the following:

- Install the _swirl_ package.
- Load the _swirl_ package.
- List the content of the environment.
- Remove everything from the environment.
- Install my _swirl_ course that is hosted on GitHub.
- With  _swirl_  you start _swirl_ and your learning experience.
  
If the course has failed to install, you can try to download the file `swirl-it.swc` from https://github.com/hubchev/swirl-it and install the course with `install\_course()`

Please choose the course _swirl-it_ and the learning module _huber-intro-1_.
You can exit _swirl_ at any time by typing `bye()` or by clicking the _Esc_ on your keyboard. 

After you have successfully finished learning module _huber-intro-1_ please go ahead with the learning module _huber-intro-2_ that is also part of my swirl course _swirl-it_.


####  _swirl_ modules on data analytical basics {.unlisted .unnumbered}

In my swirl modules _huber-data-1_, _huber-data-2_, and _huber-data-3_ I introduce some very basic statistical principles on how to analyse data.


####  _swirl_ module on the `tidyverse` package {.unlisted .unnumbered}

I compiled a short _swirl_ module to introduce the _tidyverse_ universe. This is a powerful collection of packages which I discuss later on. The learning module is also part of my _swirl-it_ course.

#### Other _swirl_ modules {.unlisted .unnumbered}

You can also install some other courses. You find a list of courses here http://swirlstats.com/scn/index.html or here https://github.com/swirldev/swirl_courses. 

I recommend this one as it gives a general overview on very basic principles of R:
```{r eval=FALSE}
library(swirl)
install_course_github("swirldev", "R_Programming_E")
swirl()
```
