---
title: "Set up R, RStudio, and R packages"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to set up R and RStudio on your machine. We will also demonstrate
  how to install R packages from CRAN, and install the tidyverse package.
---

```{r setup, include=FALSE}
library(learnr)
tutorial_options(exercise.timelimit = 60)
```



## Outline

This interactive tutorial will help you set up your computer to use R. It is for you if you need to:

* Install R on your computer
* Install the RStudio IDE
* Install R package

You can skip this tutorial if you've already done these things.

### Do you need to work through the tutorial? 

Take the quiz below to find out.

```{r quiz1, echo = FALSE}
question("Check all that you have NOT done:",
  answer("installed R on my computer", message = "* Install R"),
  answer("installed the RStudio IDE", message = "* Install RStudio IDE"),
  answer("installed the R packages for the lecture notes _How to Use R for Data Science_", message = "* Install Packages"),
  answer("None of the above. I've done them all.", correct = TRUE, message = "You can skip this tutorial!"),
  type = "multiple",
  incorrect = "This tutorial is here to help! To get set up read:"
)
```

## Install R

### How to install R

Download and install R [here](https://cloud.r-project.org/). 

-   If you are a Windows user: Click on "Download R for Windows", then click on "base", then click on the Download link.
-   If you are macOS user: Click on "Download R for (Mac) OS X", then under "Latest release:" click on R-X.X.X.pkg, where R-X.X.X is the version number. For example, the latest version of R as of March 29, 2024 was R-4.3.3.
-   If you are a Linux user: Click on "Download R for Linux" and choose your distribution for more information on installing R for your setup.

![](https://vimeo.com/203516510)

```{r letter-a, echo=FALSE}
quiz(caption = "Quiz - R",
question("What is the price you have to pay for being able to use R?",
  answer("8 € per month"),
  answer("14 € for a licence"),
  answer("0, it is open source software", correct = TRUE),
  allow_retry = TRUE
  ),
question("What sort of programming language is R?",
    answer("R is a functional programming language.", correct = TRUE, message = "R is a *functional programming language*. If you want R to do something, you need to use a function."),
    answer("R is not programming language."),
    answer("R is a language for communicating with aliens."),
    answer("R is a language for designing rocket engines."),
    allow_retry = TRUE
  ),
question("In R, everything that exists is considered an object. Which of the following statements are true regarding this concept?",
  answer("Variables holding data values are objects.", correct = TRUE),
  answer("Functions, even the user-defined ones, are objects.", correct = TRUE),
  answer("Only complex data structures like lists and data frames are objects.", correct = FALSE),
  answer("Environments used for managing scopes are not considered objects.", correct = FALSE),
  allow_retry = TRUE
)
)
```


## Install RStudio

R is a functional programming language that runs computations. RStudio is an *integrated development environment (IDE)* that provides an interface by adding many convenient features and tools. 

### How to install RStudio

Download and install RStudio [here](https://www.rstudio.com/products/rstudio/download/).

-   Scroll down to "Installers for Supported Platforms" near the bottom of the page.
-   Click on the download link corresponding to your computer's operating system.

![](https://vimeo.com/203516968)


### Test your knowledge

```{r quiz3, echo=FALSE}
quiz(caption = "Quiz - Install RStudio",
  question("What is the RStudio IDE?",
    answer("An application that makes it easier to use R.", correct = TRUE, message = "RStudio organizes your workspace and makes it easier to write, use, debug, and save R code. I highly recommend it and will rely on the RStudio IDE in several tutorials."),
    answer("An application that let's you use R without writing any code", message = "And thank goodness! Code provides a reproducible record of your work, which is essential for data science."),
    answer("A spreadsheet program like Microsoft Excel."),
    answer("Another name for R", message = "R and RStudio are two separate things. R is a language, like English. Think of RStudio as a program that helps you use the language, kind of like how a word processing program helps you write in English."),
    allow_retry = TRUE
  ),
  question("Is the RStudio IDE free to download and use?",
    answer("Yes!", correct = TRUE, message = "Like R, RStudio is free and open-source. There do exist professional versions of RStudio with enhanced features, but we will not rely on those features in these tutorials."),
    answer("No.", message = "Like R, RStudio is free and open-source. There do exist professional versions of RStudio with enhanced features, but we will not rely on those features in these tutorials.")
  )
)
```

## Install Packages

Upon successfully installing R, you gain access to functions that are part of _Base R_. This includes standard packages automatically installed and loaded with each R session, such as `stats`, `utils`, and `graphics`, providing a broad spectrum of functionalities for statistical analysis and graphical capabilities. However, the syntax in _Base R_ can become complex and less intuitive for users. Consequently, many individuals, including Hadley Wickham, the Chief Data Scientist at _Posit_ (formerly RStudio), and his team, have developed an alternative suite of packages known as the `tidyverse`. These packages share a common philosophy and syntax, emphasizing readability and ease of use. We will heavily utilize the `tidyverse` in the following sections.

### How to install R packages

To install an R package you can use the GUI of R Studio or the command line. In R Studio you can click on the *Packages* tab, then on the *Install* button, then you must search for a package and click *Install*. An alternative way to install a package is by typing

```{r, eval=FALSE}
install.packages("package_name")
```

in the console pane of RStudio and pressing Return/Enter on your keyboard. Note you must include the quotation marks around the name of the package.

![](https://vimeo.com/203516241)

Now, install frequently used packages by running the following lines of code in the Console:

```{r, eval=FALSE}
install.packages(
  c("arrow", "babynames", "curl", "duckdb", "gapminder", "ggrepel", "ggridges",
    "ggthemes", "hexbin", "janitor", "Lahman", "leaflet", "learnr", "maps",
    "nycflights13", "openxlsx", "palmerpenguins", "repurrrsive", "tidyverse",
    "tidymodels", "usethis", "WDI", "wbstats", "writexl")
  )
```

You can streamline the process using the `p_load` function as follows:

```{r, echo=T, eval=FALSE}
# Install and load the required packages using p_load
if (!require(pacman)) install.packages("pacman")

pacman::p_load(
  arrow, babynames, curl, duckdb, funcuse, gapminder, ggrepel,
  ggridges, ggthemes, hexbin, janitor, Lahman, leaflet, maps,
  nycflights13, openxlsx, palmerpenguins, repurrrsive, tidyverse,
  tidymodels, usethis, WDI, wbstats, writexl
)
```

### Test your knowledge

```{r names, echo = FALSE}
quiz(caption = "Quiz - Working with Packages",
  question("What command do you use to install packages?",
    answer("`library()`", message = "We will see what `library()` does later."),
    answer("`install.packages()`", correct = TRUE),
    answer("`install_packages()`"),
    answer("There is no command. You must visit [cran.r-project.org](http://cran.r-project.org) and download packages manually.", message = "R makes it easy to download packages. You connect to the internet and then run one of the commands above."),
    allow_retry = TRUE
  ),
  question("How often do you need to install a package on your computer?",
    answer("Every time you restart R"),
    answer("Every time you restart your computer"),
    answer("Only once. Afterwards, R can find it on your hard drive as needed.", correct = TRUE),
    answer("Never, as long as you are connected to the internet.", message = "This could be true if you are using R over a cloud service. However, if you are using R locally on your own computer, you will need to install each package that you use on your computer."),
    allow_retry = TRUE
  ),
  question("What is the tidyverse?",
    answer("A collection of packages that work well together and provide tools for common data science tasks.", correct = TRUE, message = 'You can install all of the tidyverse packages at once with `install.packages("tidyverse")`.'),
    answer("A word processing program for R"),
    answer("A code style that reads like poetry"),
    answer("An alternate universe where everyone is an R programmer."),
    allow_retry = TRUE
  )
)
```
